\chapter{Observing}

\section{Early Science Observations}

DDRAGO is currently being offered for early science observations on best-effort and shared-risk basis, on the understanding that commissioning tasks and the transient science program may take priority, and with the warning that the instrument is not fully characterized. 

The transients team currently observes by taking 60 seconds exposures and dithering randomly in a circle of diameter 1 arcmin. This seems to give a good balance between efficiency, reaching the sky limit, being able to form a sky image, and not losing too much image quality to telescope tracking errors. Other strategies are possible, and one of the aims of this shared-risk time is to allow observers to determine the best strategies for their science and to communicate their requirements and experience to us.

We aim to keep individual observing blocks to about 30 minutes of real time (i.e., 24 x 60 second exposures plus overheads). If more data is required, we recommend repeating blocks.

\section{Pointing Limits}

The telescope is located at 115.4646~{\deg} east and 31.0449~{\deg} north, and at an 
altitude of 2792 meters. The telescope has an altitude-azimuth mount and can point anywhere above a zenith distance limit of 73.4~{\deg}, which corresponds to an airmass of 3.5. However, the telescope cannot track well within 1~{\deg} of the zenith.

\section{Estimated Exposure Times}

As a rule of thumb, a typical measured 10-sigma limiting magnitudes for point sources in 60 seconds in the $g$, $r$, and $i$ filters in grey time is $\AB \approx 20$, so in $n$ exposures a typical $m$-sigma limiting magnitude is 
\begin{align}
\AB ≈ 20 + 1.25 \log_{10}(n) − 2.5 \log_{10}(m / 10)
\end{align}
These values obviously vary with seeing and sky brightness. We recommend taking initial test observations before submitting larger requests.

\section{Acknowledgements in Publications}

In publications that make use of data acquired with DDRAGO, we request that you cite these papers on the telescope and instrument at an appropriate point:

\begin{itemize}
\item \href{https://ui.adsabs.harvard.edu/abs/2022SPIE12182E..1SB/abstract}{Basa et al.\ (2022)}
\item \href{https://ui.adsabs.harvard.edu/abs/2024SPIE13096E..3DL/abstract}{Langarica et al.\ (2024)}

\end{itemize}

We also request that you include this text in the acknowledgments:

\begin{quote}
The data [or some of the data] used in this paper were acquired with the DDRAGO instrument on the COLIBRÍ telescope at the Observatorio Astronómico Nacional on the Sierra de San Pedro Mártir. COLIBRÍ and DDRAGO are funded by the Universidad Nacional Autónoma de México (CIC and DGAPA/PAPIIT IN109418, IN109224, and XXX), CONAHCyT (CF 2019/1046632 and XXX), and XXX. COLIBRÍ and DDRAGO are operated and maintained by the Observatorio Astronómico Nacional, the Instituto de Astronomía of the Universidad Nacional Autónoma de México, and XXX.
\end{quote}

There is no requirement to include members of the science operations team as authors on any publication that results from these observations in early science time or time awarded by the TACs. However, if you consider that the engineering or science operations teams have been especially helpful, you might consider mentioning them in the acknowledgements.
